---
title: "p8105_hw2_ps3070"
output: github_document
---

First, I will load the `tidyverse` and `readxl` libraries.

```{r setup, message=FALSE}

library(tidyverse)
library(readxl)
```

## PROBLEM 1

To read and clean the Mr. Trash Wheel .xlsx file, I will:
* import the data
* specify the sheet in the Excel file
* omit non-data entries
* use reasonable variable names
* omit rows that do not include dumpster-specific data
* round the number of sports balls to the nearest integer and converts the result to an integer variable (using `as.integer`)

```{r import trash wheel, message=FALSE}
data_trashwheel = read_excel(
  "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
  sheet = "Mr. Trash Wheel",
  range = "A2:N338") %>%
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls, digits = 0)))
```

Only cells from A2 to N338 are imported because column O contains notes and row 1 only contains a figure and no data.

To clean up names of the data to snake case, I used `janitor::clean_names()`.

The grand totals are included within the dataset. To omit rows that are dumpster-specific, I omitted rows that included these grand totals. Under the row `dumpster`, I dropped any rows that were recorded as `NA`/without data.


# Reading and Cleaning precipitation data

The 2018 precipitation data has empty rows, so I used `na.omit()` to avoid saving those rows.

```{r}
data_precip_2017 = read_excel(
  "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
  sheet = 4,
  range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2017)

data_precip_2018 = read_excel(
  "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
  sheet = 3,
  range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  na.omit(total)

data_precip = bind_rows(data_precip_2017, data_precip_2018) %>% 
  janitor::clean_names() %>% 
  rename(precip_in = total) %>%
  mutate(month = month.name[month])
```


